Fail Faster

### 2025-03-24 paper/paper.tex
3.7

The speed of generation becomes particularly important in
this setting. Faster generators will be able to explore more of the search 
space, produce more valid inputs to the system
under test, and uncover more bugs. Additionally,
since type derivation is a fully automatic process, the user is not required to understand 
staging in any way.

% As a result, producing values with complex preconditions is statistically unlikely.
% This state of affairs is not ideal, as inexperienced PBT users are liable to derive a generator, run their tests, find no bugs, and assume the code is correct---when in reality, most generated values are never tested at all because they do not satisfy the preconditions.
% \jwc{This paragraph reads as mostly a critique of type-deriving, and it was made already in PBT in practice...}

@@

Our implementation in OCaml uses the PPX (PreProcessor eXtension) system to synthesize 
staged generators from type definitions. However, any language that supports type-derived 
generators can implement a staged version using a similar implementation to the original. In Scala, 
for instance, the same strategy could be realized using typeclass 
resolution. In both settings, the result is a three-stage process: a 
metaprogram---via PPX or typeclasses---constructs a generator expression 
composed of staged combinators; this expression is evaluated at compile 
time to produce a specialized generator; and finally, the generator is 
executed at runtime to produce values. The entire 
pipeline remains invisible to the user. 
% \tr{TODO: Boolean list example in a figure...}
% \jwc{Unclear exactly what we want the boolean list example for here... are we going to show the type and the derived staged generator? I think that's a little too deep in the weeds for this section.}

>>>>

The speed of generation becomes particularly important in this setting. In a generator 
that produces only valid values, only a subset of them will trigger a bug; in 
a type-derived generator, however, only a subset of generated values will be valid, 
and only a subset of \textit{those} will find a bug. As a result, many more values must be produced, 
making it important to produce them as quickly as possible.

@@

Our implementation in OCaml uses the PPX (PreProcessor eXtension) system to synthesize 
staged generators from type definitions. However, any language that supports type-derived 
generators can implement a staged version using a similar implementation to the original. In Scala, 
for instance, the same strategy could be realized using typeclass 
resolution. In both settings, the result is a three-stage process: a 
metaprogram---via PPX or typeclasses---constructs a generator expression 
composed of staged combinators; this expression is evaluated at compile 
time to produce a specialized generator; and finally, the generator is 
executed at runtime to produce values.

% \tr{TODO: Boolean list example in a figure...}
% \jwc{Unclear exactly what we want the boolean list example for here... are we going to show the type and the derived staged generator? I think that's a little too deep in the weeds for this section.}
